```{r setupmutationsswiss, include=FALSE}
rm(list = ls()) ; invisible(gc()) ; set.seed(42)
library(knitr)
library(kableExtra)
if(knitr:::is_html_output()) options(knitr.table.format = "html") 
if(knitr:::is_latex_output()) options(knitr.table.format = "latex") 
library(tidyverse)
theme_set(bayesplot::theme_default())
opts_chunk$set(
  echo = F, message = F, warning = F, fig.height = 6, fig.width = 8,
  cache = T, cache.lazy = F)
```

```{r datamutationsswiss}
snv <- read_tsv("data/mutations/swiss/napoleon_mutations.tsv")
genome <- read_tsv("data/mutations/swiss/Qrob_PM1N.fa.fai", 
         col_names = c("CHROM", "length", "bytesindex", "basesperline", "bytesperline")) %>% 
  dplyr::select(CHROM, length) %>% 
  mutate(start = 0, stop = length/10^6)
```

```{r, cache=FALSE}
strelka2raw <- src_sqlite("data/mutations/swiss/strelka2_raw.sql") %>% tbl("mutations")
gatkraw <- src_sqlite("data/mutations/swiss/gatk_raw.sql") %>% tbl("mutations")
```

# Mutations swiss

This chapter describes the reanalyses of data from @Schmid-Siegert2017 currently done in the [`swiss` branch of the `detectMutations` repository](https://github.com/sylvainschmitt/detectMutations/tree/swiss).

## Mutations from Schmid-Siegert on 3P

I reported the mutations (Tab. \@ref(tab:napoTab)) from the [supplementary table 2](https://static-content.springer.com/esm/art%3A10.1038%2Fs41477-017-0066-9/MediaObjects/41477_2017_66_MOESM1_ESM.pdf) from @Schmid-Siegert2017, and aligned them back on the 3P genome (Fig. \@ref(fig:napoFig)). 
I found back only 14 of the original 17 mutations from Napoleon

```{r napoTab}
kable(snv, caption = "SNVs in the Napoleon Oak. Rerported from @Schmid-Siegert2017.")
```

```{r napoFig, fig.cap="Napoleon's original mutations on the 3P genome."}
ggplot(genome, aes(x = CHROM, xend = CHROM)) +
  geom_segment(aes(y = start, yend = stop), size = 3, col = "grey") +
  geom_point(aes(y = POS/10^6), col = "red", data = snv) +
  ggrepel::geom_label_repel(aes(y = POS/10^6, label = Mutation), col = "red", data = snv) +
  coord_flip() +
  ylab("Position (Mb)") +
  theme(axis.line.y = element_blank(), axis.title.y = element_blank(), 
        axis.ticks.y = element_blank())
```
## `Strelka2`

`Strelka2` produced `r round(collect(tally(strelka2raw))$n/10^6, 1)` millions of candidate mutations.

### Overlap with mutations from Schmid-Siegert

I tried to find back Napoleon's original mutations to have a look to their metrics.
I found back only 12 out of the 14 expected mutations (86%) (Tab. \@ref(tab:napoMutTab) and Fig. \@ref(fig:napoMutFig)).
**Beware, `Strelka2` is detecting putative mutations in the normal sample !**
I looked at different metrics for each  (Fig. \@ref(fig:overlapMetrics)):

* `mutation_DP` and `normal_DP` are the read depth for the two sample, and shows as expected values between half and two times the mean coverage (60X)
* `normal_altCountT1` is the number of alternate allele count in the normal sample, should be 0, but is equal to 3 and 4 (9% of reads) for two SNVs
* `mutation_altCountT1` is the number of alternate allele count in the mutated sample, should be not too low, and is most the time above 5

**The main conclusion is that the mutations detected by @Schmid-Siegert2017 have not always no reads in the "normal" sample and that they show a wide variation of allelic frequency.**

```{r napoMut}
# overlapS <- lapply(1:nrow(snv), function(i)
#   strelka2raw %>%
#     filter(CHROM == local(snv[i,]$CHROM)) %>%
#     filter(POS == local(snv[i,]$POS)) %>%
#     collect() %>%
#     mutate(Mutation  = snv[i,]$Mutation)
# ) %>% bind_rows()
# write_tsv(overlapS, file = "save/overlap_strelka2.tsv")
overlapS <- read_tsv("save/overlap_strelka2.tsv")
overlapSfiltered <- overlapS %>% 
  group_by(Mutation) %>% 
  filter(normal_altCountT1 == min(normal_altCountT1))
```

```{r napoMutTab}
overlapS %>% 
  dplyr::select(Mutation, tumor, normal, REF, ALT, mutation_altCountT1, mutation_refCountT1,
                normal_altCountT1, normal_refCountT1, mutation_AF) %>% 
  kable(caption = "Overlap between candidate mutations and Napoleon's original mutations.", 
        col.names = c("Mutation", "Mutated", "Normal", "Ref", "Alt", 
                      "Mutated\nAltCount", "Mutated\nRefCount", "Normal\nAltCount", "Normal\nRefCount", "Allelic fraction"))
```

```{r napoMutFig, fig.cap="Overlap between candidate mutations and Napoleon's original mutations: allele frequency (A) and positions on the 3P genome (B)."}
g1 <- ggplot(overlapSfiltered, aes(mutation_AF)) +
  geom_histogram() +
  ggtitle("", paste("N =", nrow(overlapSfiltered))) +
  xlab("Allele frequency")
g2 <- ggplot(genome, aes(x = CHROM, xend = CHROM)) +
  geom_segment(aes(y = start, yend = stop), size = 3, col = "lightgrey") +
  geom_point(aes(y = POS/10^6, col = mutation_AF), data = overlapSfiltered, size = 1) +
  ggrepel::geom_label_repel(aes(y = POS/10^6, label = Mutation), col = "red", 
                            data = dplyr::select(overlapSfiltered, CHROM, POS, Mutation) %>% unique()) +
  coord_flip() +
  ylab("Position (Mb)") +
  theme(axis.line.y = element_blank(), axis.title.y = element_blank(), axis.ticks.y = element_blank()) +
  viridis::scale_color_viridis(option = "inferno")
cowplot::plot_grid(g1, g2, nrow = 2, labels = c("A", "B"))
```

```{r overlapMetrics, fig.cap="Evaluation of the overlap between candidate mutations and Napoleon's original mutations."}
overlapSfiltered %>% 
  group_by(Mutation) %>% 
  filter(normal_altCountT1 == min(normal_altCountT1)) %>% 
  dplyr::select(Mutation, FILTER, mutation_DP, normal_DP,
                mutation_altCountT1, normal_altCountT1, QSS) %>% 
  reshape2::melt(c("Mutation", "FILTER")) %>% 
  ggplot(aes(value, fill = FILTER)) +
  geom_histogram() +
  facet_wrap(~ variable, scales = "free") +
  theme(legend.position = c(0.8, 0.2)) +
  geom_vline(aes(xintercept = value, fill = NA), size = 1.1, col = "black", linetype = "dashed",
             data = data.frame(
               value = c(30-0.5, 60*1.1+0.5, 30-0.5, 60*1.1+0.5, 10-0.5, 0+0.5, 20),
               variable = c("mutation_DP", "mutation_DP", 
                            "normal_DP", "normal_DP", 
                            "mutation_altCountT1", "normal_altCountT1", "QSS")
             ))
```

### Filtering

We originally planned to filter mutations with following filters:

* A read depth for the two sample between half and two times the mean coverage (`normal_DP <= 120, normal_DP >= 30, mutation_DP <= 120, mutation_DP >= 30`)
* A null number of alternate allele count in the normal sample (`normal_altCountT1 == 0`)
* A minimum of 10 alternate allele count in the mutated sample (`mutation_altCountT1 >= 10`)
* An allele frequency of mutations below 0.5 (`mutation_AF <= 0.5`)

We obtained 228 candidates  (Fig \@ref(fig:mutFilteredS)).
We then used the suggested automatic filter of `Strelka2`,
resulting in a robust dataset of 91 mutations (Fig \@ref(fig:mutRobustS)).

```{r mutS}
mutFilteredS <- strelka2raw %>% 
  filter(normal_DP <= 120, normal_DP >= 30) %>% 
  filter(mutation_DP <= 120, mutation_DP >= 30) %>% 
  filter(normal_altCountT1 == 0) %>% 
  filter(mutation_altCountT1 >= 10) %>% 
  filter(mutation_altCountT1 < 30) %>% 
  collect() 
mutRobustS <- filter(mutFilteredS, FILTER == "PASS")
```

```{r mutFilteredS, fig.cap="Filtered mutations allele frequency (A) and positions on the 3P genome (B)."}
g1 <- ggplot(mutFilteredS, aes(mutation_AF)) +
  geom_histogram() +
  ggtitle("", paste("N =", nrow(mutFilteredS))) +
  xlab("Allele frequency")
g2 <- ggplot(genome, aes(x = CHROM, xend = CHROM)) +
  geom_segment(aes(y = start, yend = stop), size = 3, col = "lightgrey") +
  geom_point(aes(y = POS/10^6, col = mutation_AF), data = mutFilteredS, size = 1) +
  coord_flip() +
  ylab("Position (Mb)") +
  theme(axis.line.y = element_blank(), axis.title.y = element_blank(), axis.ticks.y = element_blank()) +
  viridis::scale_color_viridis(option = "inferno")
cowplot::plot_grid(g1, g2, nrow = 2, labels = c("A", "B"))
```

```{r mutRobustS, fig.cap="Robust mutations allele frequency (A) and positions on the 3P genome (B)."}
g1 <- ggplot(mutRobustS, aes(mutation_AF)) +
  geom_histogram() +
  ggtitle("", paste("N =", nrow(mutRobustS))) +
  xlab("Allele frequency")
g2 <- ggplot(genome, aes(x = CHROM, xend = CHROM)) +
  geom_segment(aes(y = start, yend = stop), size = 3, col = "lightgrey") +
  geom_point(aes(y = POS/10^6, col = mutation_AF), data = mutRobustS, size = 1) +
  coord_flip() +
  ylab("Position (Mb)") +
  theme(axis.line.y = element_blank(), axis.title.y = element_blank(), axis.ticks.y = element_blank()) +
  viridis::scale_color_viridis(option = "inferno")
cowplot::plot_grid(g1, g2, nrow = 2, labels = c("A", "B"))
```

## `GATK`

<!-- `GATK` produced 49.8 millions of candidates! -->

### Overlap with mutations from Schmid-Siegert

<!-- I tried to find back Napoleon's original mutations to have a look to their metrics. -->
<!-- I found back only 8 out of the 11 expected mutations (73%) (Tab. \@ref(tab:napoMutTab2) and Fig. \@ref(fig:napoMutFig2)). -->
<!-- I looked at different metrics for each  (Fig. \@ref(fig:overlapMetrics2)): -->

<!-- * `mutation_DP` and `normal_DP` are the read depth for the two sample, and shows as expected values between half and two times the mean coverage (60X) -->
<!-- * `normal_altCountT1` is the number of alternate allele count in the normal sample, should be 0, but is equal to 3 and 4 (9% of reads) for two SNVs -->
<!-- * `mutation_altCountT1` is the number of alternate allele count in the mutated sample, should be not too low, and is most the time above 5 -->

<!-- **The main conclusion is that the mutations detected by @Schmid-Siegert2017 have not always no reads in the "normal" sample and that they show a wide variation of allelic frequency. For instance, the SNV6 not detected by `Strelka2` has 10 copies in the lower branch!** -->

```{r napoMut2}
# overlapG <- lapply(1:nrow(snv), function(i)
#   gatkraw %>%
#     filter(CHROM == local(snv[i,]$CHROM)) %>%
#     filter(POS == local(snv[i,]$POS)) %>%
#     collect() %>%
#     mutate(Mutation  = snv[i,]$Mutation)
# ) %>% bind_rows()
# overlapG <- mutate(overlapG, AF = altCount_tumor/(altCount_tumor+refCount_tumor))
# write_tsv(overlapG, file = "save/overlap_gatk.tsv")
# overlapG <- read_tsv("save/overlap_gatk.tsv") 
```

```{r napoMutTab2}
# overlapG %>% 
#   dplyr::select(Mutation, tumor, normal, REF, ALT, altCount_tumor, refCount_tumor, altCount_normal, refCount_normal, AF) %>% 
#   kable(caption = "Overlap between candidate mutations and Napoleon's original mutations.", 
#         col.names = c("Mutation", "Mutated", "Normal", "Ref", "Alt", 
#                       "Mutated\nAltCount", "Mutated\nRefCount", "Normal\nAltCount", "Normal\nRefCount", "Allelic fraction")) 
```

```{r napoMutFig2, fig.cap="Overlap between candidate mutations and Napoleon's original mutations: allele frequency (A) and positions on the 3P genome (B)."}
# g1 <- ggplot(overlapG, aes(AF)) +
#   geom_histogram() +
#   ggtitle("", paste("N =", nrow(overlapG))) +
#   xlab("Allele frequency")
# g2 <- ggplot(genome, aes(x = CHROM, xend = CHROM)) +
#   geom_segment(aes(y = start, yend = stop), size = 3, col = "lightgrey") +
#   geom_point(aes(y = POS/10^6, col = AF), data = overlapG, size = 1) +
#   ggrepel::geom_label_repel(aes(y = POS/10^6, label = Mutation), col = "red", 
#                             data = dplyr::select(overlapG, CHROM, POS, Mutation) %>% unique()) +
#   coord_flip() +
#   ylab("Position (Mb)") +
#   theme(axis.line.y = element_blank(), axis.title.y = element_blank(), axis.ticks.y = element_blank()) +
#   viridis::scale_color_viridis(option = "inferno")
# cowplot::plot_grid(g1, g2, nrow = 2, labels = c("A", "B")) 
```

```{r overlapMetrics2, fig.cap="Evaluation of the overlap between candidate mutations and Napoleon's original mutations."}
# overlapG %>% 
#   group_by(Mutation) %>% 
#   dplyr::select(Mutation, DP_tumor, DP_normal, altCount_tumor, altCount_normal) %>% 
#   reshape2::melt(c("Mutation")) %>% 
#   ggplot(aes(value)) +
#   geom_histogram() +
#   facet_wrap(~ variable, scales = "free") +
#   geom_vline(aes(xintercept = value), size = 1.1, col = "black", linetype = "dashed",
#              data = data.frame(
#                value = c(30-0.5, 60*1.1+0.5, 30-0.5, 60*1.1+0.5, 10-0.5, 0+0.5),
#                variable = c("DP_tumor", "DP_tumor", 
#                             "DP_normal", "DP_normal", 
#                             "altCount_tumor", "altCount_normal")
#              )) 
```

### Filtering

<!-- We couldn't use the read depth for the normal sample and the quality score above not reported in `GATK`. -->
<!-- We first filtered mutations with available parameters: -->

<!-- * A read depth for the mutated sample between half and two times the mean coverage (`DP_tumor <= 60*1.1, DP_tumor >= 30`) -->
<!-- * A null number of alternate allele count in the normal sample (`normal_altCountT1 == 0`) -->
<!-- * A minimum of 10 alternate allele count in the mutated sample (`altCount_tumor >= 10`) -->
<!-- * An allele frequency of mutations below 0.5 (`mutation_AF <= 0.5`) -->

<!-- But we obtained 1.9 million candidates! -->
<!-- We thus looked for the overlap between `GATK` candidates and the suggested automatic filter of `Strelka2`, -->
<!-- resulting in a robust dataset of 47 mutations (Fig \@ref(fig:mutRobustG)). -->
<!-- Nevertheless, for comparisons we want automatic filtering based on simple metrics. -->
<!-- I thus compared the distribution of the robust and raw mutations for different metrics alone and pairwise (Fig. \@ref(fig:mutMetricsG)). -->
<!-- And unfortunately, it does not seem that the metrics offered by `GATK` are sufficient to find back robust mutations. -->
<!-- They do show differences, e.g. for `MQRankSum`, but resulting in the loss of numerous robust mutations and still too much candidates. -->

<!-- For instance, we could compare 4 datasets overlaps and distribution in allelic frequency: -->

<!-- * mutations from Schmid-Siegert on the 12 first chromosomes from 3P -->
<!-- * mutations from `GATK` with weak filtering -->
<!-- * mutations from `Strelka2` with weak filtering -->
<!-- * mutations from `Strelka2` with robust filtering -->

```{r mutRawG}
# mutRawG <- gatkraw %>% 
#   mutate(AF = altCount_tumor/(altCount_tumor+refCount_tumor)) %>% 
#   filter(DP_tumor <= 60*1.1, DP_tumor >= 30) %>% 
#   filter(altCount_normal == 0) %>% 
#   filter(altCount_tumor >= 10) %>% 
#   filter(AF <= 0.5) %>% 
#   collect() %>% 
#   left_join(dplyr::select(mutRobust, CHROM, POS) %>% 
#               mutate(Robust = 1)) %>% 
#   mutate(Robust = as.factor(ifelse(is.na(Robust), 0, 1)))
# mutRobustG <- filter(mutRawG, Robust == 1) 
```

```{r mutRobustG, fig.cap="Robust mutations allele frequency (A) and positions on the 3P genome (B)."}
# g1 <- ggplot(mutRobustG, aes(AF)) +
#   geom_histogram() +
#   ggtitle("", paste("N =", nrow(mutRobustG))) +
#   xlab("Allele frequency")
# g2 <- ggplot(genome, aes(x = CHROM, xend = CHROM)) +
#   geom_segment(aes(y = start, yend = stop), size = 3, col = "lightgrey") +
#   geom_point(aes(y = POS/10^6, col = AF), data = mutRobustG, size = 1) +
#   coord_flip() +
#   ylab("Position (Mb)") +
#   theme(axis.line.y = element_blank(), axis.title.y = element_blank(), axis.ticks.y = element_blank()) +
#   viridis::scale_color_viridis(option = "inferno")
# cowplot::plot_grid(g1, g2, nrow = 2, labels = c("A", "B")) 
```

```{r mutMetrics2, fig.cap="Different metrics single and pair plots for raw and robust mutations."}
# mutRawG %>% 
#   dplyr::select(Robust, DP_tumor, altCount_tumor, QUAL, BaseQRankSum, FS, MQ, MQRankSum, QD, ReadPosRankSum, SOR, AF) %>% 
#   arrange(desc(Robust)) %>% 
#   head(10^4) %>% 
#   reshape2::melt("Robust") %>% 
#   ggplot(aes(value, col = Robust), fill = NA) +
#   geom_density() +
#   facet_wrap(~variable, scales = "free") 
```

## Conclusion

<!-- I looked at 6 datasets overlaps and distribution in allelic frequency: -->

<!-- * mutations from Schmid-Siegert on the 4 first chromosomes from 3P -->
<!-- * mutations from `GATK` with **low** filtering on the 4 first chromosomes from 3P -->
<!-- * mutations from `Strelka2` with **low** filtering on the 4 first chromosomes from 3P -->
<!-- * mutations from `GATK` with **weak** filtering on the 4 first chromosomes from 3P -->
<!-- * mutations from `Strelka2` with **weak** filtering on the 4 first chromosomes from 3P -->
<!-- * mutations from `Strelka2` with **robust** filtering on the 4 first chromosomes from 3P -->

<!-- The filtering are: -->

<!-- * **Low** filtering: -->
<!--     * A read depth for the mutated sample between 50% and 110% of the expected coverage -->
<!--     * A null number of alternate allele count in the normal sample -->
<!--     * A minimum of 5 alternate allele count in the mutated sample -->
<!--     * An allele frequency of mutations below 0.5 -->
<!-- * **Weak** filtering: -->
<!--     * A read depth for the mutated sample between 50% and 110% of the expected coverage -->
<!--     * A null number of alternate allele count in the normal sample -->
<!--     * A minimum of 10 alternate allele count in the mutated sample -->
<!--     * An allele frequency of mutations below 0.5 -->
<!-- * **Robust** filtering: -->
<!--     * A read depth for the two sample between 50% and 110% of the expected coverage (`normal_DP < 60*1.1, normal_DP >= 30, mutation_DP < 60*1.1, mutation_DP >= 30`) -->
<!--     * A null number of alternate allele count in the normal sample (`normal_altCountT1 == 0`) -->
<!--     * A minimum of 10 alternate allele count in the mutated sample (`mutation_altCountT1 >= 10`) -->
<!--     * An allele frequency of mutations below 0.5 (`mutation_AF <= 0.5`) -->
<!--     * A quality score above 20 (`QSS > 20`) -->
<!--     * The suggested automatic filter of `Strelka2` -->

```{r datasetsswiss}
# schmid <- bind_rows(overlap, overlapG) %>%
#   dplyr::select(CHROM, POS, normal_AF, AF) %>%
#   mutate(AF = ifelse(is.na(AF), normal_AF, AF)) %>%
#   dplyr::select(-normal_AF) %>%
#   mutate(SNV = paste0(CHROM, "_", "Pos", POS)) %>%
#   group_by(CHROM, POS, SNV) %>%
#   summarise(AF = max(AF))
# gatk_w <- gatkraw %>%
#   mutate(AF = altCount_tumor/(altCount_tumor+refCount_tumor)) %>%
#   filter(DP_tumor <= 60*1.1, DP_tumor >= 30) %>%
#   filter(altCount_normal == 0) %>%
#   filter(altCount_tumor >= 10) %>%
#   filter(AF <= 0.5) %>%
#   mutate(SNV = paste0(CHROM, "_", "Pos", as.integer(POS))) %>%
#   dplyr::select(CHROM, POS, SNV, AF) %>%
#   collect()
# strelka_w <- strelka2raw %>% 
#   filter(mutation_DP <= 60*1.1, mutation_DP >= 30) %>% 
#   filter(normal_altCountT1 == 0) %>% 
#   filter(mutation_altCountT1 >= 10) %>% 
#   filter(mutation_AF <= 0.5) %>% 
#   mutate(SNV = paste0(CHROM, "_", "Pos", as.integer(POS))) %>% 
#   rename(AF = mutation_AF) %>% 
#   dplyr::select(CHROM, POS, SNV, AF) %>% 
#   collect()
# strelka_r <- strelka2raw %>% 
#   filter(normal_DP <= 60*1.1, normal_DP >= 30) %>% 
#   filter(mutation_DP <= 60*1.1, mutation_DP >= 30) %>% 
#   filter(normal_altCountT1 == 0) %>% 
#   filter(mutation_altCountT1 >= 10) %>% 
#   filter(mutation_AF <= 0.5) %>% 
#   filter(mutation_altCountT1 < 30) %>% 
#   filter(QSS > 20) %>% 
#   filter(FILTER == "PASS") %>% 
#   mutate(SNV = paste0(CHROM, "_", "Pos", as.integer(POS))) %>% 
#   rename(AF = mutation_AF) %>% 
#   dplyr::select(CHROM, POS, SNV, AF) %>% 
#   collect() 
# datasets <- list(
#   "Schmid" = schmid,
#   "GATK" = gatk_w,
#   "Strelka2" = strelka_w,
#   "Robust" = strelka_r
# )
datasets <- list(
  "Schmid" = overlapSfiltered %>% 
    mutate(SNV = paste0(CHROM, "_", "Pos", as.integer(POS))) %>%
    rename(AF = mutation_AF) %>%
    dplyr::select(CHROM, POS, SNV, AF),
  # "GATK" = gatk_w,
  "Strelka2" = mutFilteredS %>% 
    mutate(SNV = paste0(CHROM, "_", "Pos", as.integer(POS))) %>%
    rename(AF = mutation_AF) %>%
    dplyr::select(CHROM, POS, SNV, AF),
  "Robust" = mutRobustS %>% 
    mutate(SNV = paste0(CHROM, "_", "Pos", as.integer(POS))) %>%
    rename(AF = mutation_AF) %>%
    dplyr::select(CHROM, POS, SNV, AF)
)
# rm(schmid, gatk_w, strelka_w, strelka_r)
```

```{r datasetsswissTab}
# lapply(datasets, nrow) %>% 
#   bind_rows(.id = "dataset") %>% 
#   reshape2::melt(variable.name = "Dataset", value.name = "Number of candidates") %>% 
#   kable(caption = "Size of the different datasets.", format.args = list(big.mark = " ")) 
```

```{r datasetsswissAF, fig.cap="Allele frequency for the different datasets."}
bind_rows(datasets, .id = "dataset") %>%
  ggplot(aes(AF, col = dataset), fill = NA) +
  geom_density() +
  scale_color_discrete("Dataset") +
  xlab("Allelic frequency")
```

```{r datasetsswissVenn2}
ggvenn::ggvenn(lapply(datasets, `[[`, "SNV"), show_percentage = F)
```

```{r datasetsswissVenn, eval=F}
# overlaps <- lapply(datasets, select, SNV, AF) %>% 
#   bind_rows(.id = "dataset") %>% 
#   ungroup() %>% 
#   select(dataset, SNV) %>% 
#   mutate(value = "1") %>% 
#   reshape2::dcast(SNV ~ dataset) %>% 
#   mutate_all(funs(ifelse(is.na(.), 0, .))) %>% 
#   select(-SNV) %>% 
#   mutate_all(as.numeric) %>% 
#   as.matrix() %>% 
#   crossprod() %>% 
#   as.data.frame() %>% 
#   rownames_to_column("dataset")
# write_tsv(overlaps, file = "save/overlaps.tsv")
# read_tsv("save/overlaps.tsv") %>% 
#   kable(caption = "Overlaps among the different datasets.", format.args = list(big.mark = " "))
```
