```{r setupAbibliometry, include=FALSE}
rm(list = ls()) ; invisible(gc()) ; set.seed(42)
library(knitr)
library(tidyverse)
library(bibliometrix)  
file <- "data/biblio/Dimensions-Publication-2021-04-08_12-49-49.csv"
biblio <- convert2df(file, dbsource = "dimensions", format = "csv")
theme_set(bayesplot::theme_default())
opts_chunk$set(
  echo = F, message = F, warning = F, fig.height = 6, fig.width = 8,
  cache = T, cache.lazy = F)
```

# Appendix 4: Bibliometry

This appendix help to explore automtically the litterature with the package `bibliometrix` and [dimension](https://app.dimensions.ai) collection.

I used keywords from few selected papers to design the following query:

> intraorganismal selection, genomic mutation, life history, genetic variation, flowering plant, Article or Preprint

And I then manually curated the 25 obtained articles to articles related to our subject, resulting in only **16** articles.

## Description

At the exception of `r biblio[biblio$PY==1997,]$SR`, this is a relatively new field (Fig \@ref(fig:plotRes)) that emerged in 2011 and that started peaking in 2018.

```{r load, include=F, eval=F}
results <- biblioAnalysis(biblio, sep = ";")
sum.res <- summary(results)
plot.res <- plot(x = results, k = 10, pause = FALSE)
```

```{r sumRes, eval=F}
kable(sum.res$MainInformationDF, caption ="Main informations.")
```

```{r plotRes, fig.width=10, fig.cap="Summary plot.", eval=F}
cowplot::plot_grid(plot.res$AnnualScientProd,
                   plot.res$AverArtCitperYear,
                   plot.res$MostProdAuthors,
                   plot.res$MostProdCountries)
```

## Citation

```{r, eval=F}
ggplot(biblio, aes(SR, TC)) + 
  geom_col() +
  coord_flip() +
  xlab("") + ylab("Time cited")
```


## Cited references

Cited references should be further check as they include articles of primary interest such as the first:

> Folse, H.J., III and Roughgarden, J. (2012), DIRECT BENEFITS OF GENETIC MOSAICISM AND INTRAORGANISMAL SELECTION: MODELING COEVOLUTION BETWEEN A LONG‐LIVED TREE AND A SHORT‐LIVED HERBIVORE. Evolution, 66: 1091-1113. https://doi.org/10.1111/j.1558-5646.2011.01500.x

```{r cr, eval=F}
CR <- citations(biblio, field = "article", sep = ";")
kable(cbind(CR$Cited[1:20]), caption = "20 most cited articles in the bibliography.")
```

## Co-citation

```{r cocitart, fig.cap = "Top co citation articles.", eval=F}
NetMatrix <- biblioNetwork(biblio, analysis = "co-citation", network = "references", sep = ";")
net=networkPlot(NetMatrix, n = 20, 
                type = "fruchterman", size.cex=TRUE, size=20, remove.multiple=FALSE, labelsize=0.7,edgesize = 10, edges.min=5)
```

```{r cocitjournals, fig.cap = "Top co citation journals.", eval=F}
M=metaTagExtraction(biblio,"CR_SO",sep=";")
NetMatrix <- biblioNetwork(M, analysis = "co-citation", network = "sources", sep = ";")
net=networkPlot(NetMatrix, n = 20,
                type = "auto", size.cex=TRUE, size=15, remove.multiple=FALSE, labelsize=0.7,edgesize = 10, edges.min=5)
```

